Luminance constancy under fixed geometry

### 2018-05-31 Draft5/draft5.tex
small edits

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
	\centering
	\caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_a.png}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.55 \textwidth}   
        \caption{}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_b.pdf}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{{\bf Color and luminance constancy:} (a)  The light reflected from an object to the eye depends both on the surface reflectance of the object and on the illumination of the object. 
The reflected light also depends on geometric factors, such as the object's shape, pose, and position. 
The human visual system accounts for variations in the reflected light due to object-extrinsic factors and produces a percept of object color that is relatively stable across such variations, an ability called color constancy. 
(b) Images of a sphere under a fixed illuminant.  
Down each column, the sphere's light reflectance value (LRV) varies, but the relative reflectance spectrum (i.e. its spectral shape) is held fixed.
Across each row, the shape of the reflectance spectrum varies, but the LRV is held fixed.
We cast the problem of computational luminance constancy to be that of estimating the LRV of a target object from the image, across variation in other scene factors. 
Specifically, we study variation in target object relative surface reflectance, variation in illumination spectrum, and variation in the surface reflectance of the other objects in the scene.}
 \end{figure}
%% DHB: Need to decide and then change with search through document: "relative reflectance spectrum" or "shape of reflectance spectrum".
% I prefer relative, because "shape" in a vision paper generally refers to object shape.
% And we found at least one instance where trying to use "shape" led to an awkward construction.

@@

\section*{Methods} \label{Methods}
\subsection{Overview}
There are four key parts to our methods.  First, we generate a labeled set of training images.  Second, we use a model of the early visual system to compute the responses of the cone photoreceptor mosaic to the labeled images. Third, we apply AMA to the cone responses, as well as to a contrast normalized version of the cone responses, to learn task-optimal receptive fields (RFs). Fourth, we evaluate how well the responses of the RFs may be decoded to achieve luminance constancy.

% Figure 2
\begin{figure}
\centering
\begin{subfigure}[b]{0.22 \textwidth}
        \caption{Library }
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_a.pdf}
        \label{fig:baseSceneLibrary}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \caption{Mill}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_b.pdf}
        \label{fig:baseSceneMill}
    \end{subfigure}    
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \caption{Table-Chairs}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_c.pdf}
        \label{fig:baseSceneTableChairs}
    \end{subfigure}
    \caption{{\bf Base scenes.} Each subplot shows a rendering of one base scene without additional inserted objects.  The reflectance spectra of the distinct surfaces and the spectral power distributions of the light sources in each scene have been assigned randomly from statistical models of naturally occurring spectra (see text). The images in this figure have been scaled and tone mapped individually for illustrative purposes.}
\label{fig:baseScenes}. \end{figure}
%% DHB: Let's get the scaling and tone mapping used for these under control, so we know what it is.
%% DHB: Maybe also say that the illumination scaling is much smaller than the range used in the study, so that it is possible to see it with common scaling.

@@

% Figure 4
\begin{figure}
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \caption{Target position}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_a.pdf}
        \label{fig:targetPositionVariation}
    \end{subfigure}
    ~
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \caption{Target size/Orientation}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_b.pdf}
        \label{fig:targetSizeOrientation}
    \end{subfigure}
~
\centering
	\begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \caption{Target spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_c.pdf}
        \label{fig:targetVariation}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \caption{Illumination spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_d.pdf}
        \label{fig:illuminationVariation}
    \end{subfigure}
          ~  
    \begin{subfigure}[b]{0.18 \textwidth}
    \centering
        \caption{Background spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_e.pdf}
        \label{fig:backGroundVariation}
    \end{subfigure}

@@

% Figure 5: Conditions Studied
\begin{figure}
\centering
	\begin{subfigure}[b]{0.33 \textwidth}
		\caption{Condition 1}
%		\includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_a.png}
		\includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_b.pdf}
 		\label{fig:backgroundVarying}
	\end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
        \caption{Condition 2}	
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure5/Figure5_b.pdf}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}[b]{0.33 \textwidth}
	\caption{Condition 3}	
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure5/Figure5_c.pdf}        
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Dataset Conditions 1-3:} sRGB renderings of illustrative images from the datasets for Conditions 1, 2, and 3. Five images are shown for each light reflectance value (LRV). We studied three types of spectral variations. (a) Condition 1: variable target object relative reflectance spectrum, fixed light source spectra, fixed background object spectra (same as Fig.~\ref{fig:introExampleFigure}). 
(b) Condition 2: variable target object relative reflectance spectrum, variable light source spectra, fixed background object spectra. (c) Condition 3: variable target object relative reflectance spectrum, variable light source spectra, variable background object spectra. The spheres in each row of each panel have the same LRV, while the spheres in each column of each panel have the same relative surface reflectance.  Across panels a-c, spheres in corresponding locations have the same surface reflectance. In all three panels, the overall light source spectra scale factors (see \nameref{Methods}) were drawn from a uniform distribution on the range [0.5, 1.5]. This is smaller than the variation we studied, but allows us to show the variation within the dynamic range available for display in publication. The sRGB values for all three subplots were normalized using a common scale factor prior to gamma correction.} 
\label{fig:studiedCases}
\end{figure}

For the primary results, we used the ``{\it Library}'' base scene and a spherical target object.
The library base scene contains 2 area lights. 
We inserted one additional spherical light source into the scene.
The position and size of the inserted object, inserted light source and viewpoint were held fixed across all 
scenes in the database. The geometry is shown in Figure~\ref{fig:3DScene}.
Surface and illuminant spectra were drawn according to statistical models of naturally occurring spectra, as described below.
Multispectral images ($320 \times 240$ pixels) were rendered at 31 evenly-spaced wavelengths between $400$nm and $700$nm.
\subsubsection{Illumination spectra}
% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{Granada dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Statistical model of illumination spectra:} (a) Normalized Granada dataset. Each spectrum is normalized by its mean power. (b) Sample spectra generated using the statistical model derived from normalized Granada dataset. (c) CIE xy chromaticities of the Granda dataset (black) and the samples shown in b (red). (d) sRGB renditions of the samples shown in b.}
\label{fig:illuminant}
\end{figure}

To generate illumination spectra, we first developed a statistical model of the Granada daylight measurements (\citeNP{hernandez2001color}; \href{http://colorimaginglab.ugr.es/pages/Data}{http://colorimaginglab.ugr.es/pages/Data}) and then sampled randomly from the model (see appendix).
To match the wavelength spacing we use in rendering, we resampled the wavelength spacing of the Granada spectra to
to 31 evenly-spaced wavelengths between 400 nm and 700 nm.

The overall intensity of the Granada spectra spans approximately three orders of magnitude.
Figure~\ref{fig:granadaData} plots the measurements in normalized form to make apparent the variation in the relative spectra.
To normalize the spectra, we divided each spectrum by its mean power.

@@

We focussed our analysis on image regions local to the target by cropping the rendered images to $1 \times 1$ degrees of visual angle around the target object ($51 \times 51$ pixels; Figure~\ref{fig:croppedImage}). In foveal primary visual cortex, receptive fields have a spatial extent of approximately 1 degree of visual angle (REFS). 
%% VS: Needed a reference from Johannes.
The local analysis is thus motivated by the fact that neural receptive fields early in the visual pathways (e.g., retina, primary visual cortex) pool information locally (REFS).
%% DHB: Johannes to provide his favorite citation or two.  Some review.

@@

\subsubsection*{Learning optimal filters}
AMA is a task-specific Bayesian method for dimensionality reduction.
When provided with a labeled training set, a receptive field response model, a decoder that uses these responses to estimate the stimulus label, and a cost function, AMA returns a set of N linear receptive fields, where N is a parameter of the procedure.
These receptive fields are chosen so that decoding their responses leads to minimum-expected-cost estimation, relative to any other choice of N linear receptive fields.
This is possible because given a training set together with a set of N linear receptive fields and an explicit cost function, it is possible to determine the minimum-expected-cost estimator that maps the receptive field responses to the stimulus labels, on the assumption that all images in the training set are equally likely.
Moreover, the corresponding expected estimation cost may be explicitly computed.
Thus, AMA searches over the space of linear receptive fields to find the set that minimizes the expected estimation cost.
We used the Kullback-Leibler divergence cost function (corresponding to the maximum a posteriori estimator) and means squared cost function (corresponding to mean estimator) and assumed that receptive field responses were corrupted by scaled Gaussian noise (i.e. Poisson-like noise with a fano factor of 1.3).
Both, the Kullback-Leibler divergence cost function and mean squared cost function gave similar results.
The results in this work are for the Kullback-Leibler divergence cost function (see below).
Details of how AMA learns filters and how the filter responses are optimally decoded are provided in previously published work \cite{geisler2009optimal,burge2017accuracy,jaini2017linking}.
For each condition, we learned the filters with a training set consisting of 90\% of the images.

@@

Once the optimal receptive fields have been learned, we need to develop a general decoder that can be applied to arbitrary test stimuli.
A general decoder is necessary because the decoder used to learn the receptive fields requires the response mean and response variance of the filters to every labelled stimulus in the training set \cite{geisler2009optimal,burge2017accuracy}.
To proceed, first we use the AMA optimal receptive fields and the training dataset to find the conditional distributions of the receptive field responses, conditioned on the labelled target object LRVs.
Then, we approximate these with multivariate Gaussian distributions.
To estimate the LRV of the target object in a test image, we use Bayes' rule together with the Gaussian conditional distributions and a uniform prior to obtain the posterior distribution over LRV values given the filter responses.
The optimal estimate is the LRV value that minimizes the expected value of the cost function, where the expectation is taken over the posterior.
Here we again used both Kullback-Leibler divergence cost function and mean square cost function, so that the optimal estimator is the maximum a posteriori estimate and the mean estimate respectively.
The results in both cases were comparable.
We present the results for the Kullback-Leibler divergence cost function case as the estimates obtained by the mean squared cost function have a bias near the lower and the upper limits of the LRV values used in our simulations. 
The bias results because the LRV values are bounded in the range [0.2, 0.6] in estimating the posterior.

The Gaussian approximations for the posterior were estimated using the same training set that was used to learn the receptive fields.
Estimation performance was evaluated using a test set consisting of the 10\% of images that were not in the training set.
For our primary results, estimation was based on 6 receptive fields.
%% JDB: I thought we were using a squared error cost function. See above?
%% DHB: Wouldn't it make more sense to use MSE and posterior mean for both learning and decoding, since we are reporting relative RMSE as our performance measure?
%% VS: The results are similar for both cost function. MSE estimates are biased at the ends.

\subsubsection*{Baseline methods}

To provide baselines for evaluation of the estimation method described above, we used linear regression.
First, we solved for the weights on the average L, M and S cone excitations corresponding to the target object that best predicted the target LRV values.
We took the average cone responses from a $3 \times 3$ region at the center of the target object.
This baseline method disregards information carried by the image regions corresponding to the background objects.
We also performed regression on the contrast normalized version of the cone responses.
This second baseline method indirectly incorporates information carried in the image regions corresponding to the background objects,
through their effect on the contrast normalized responses at the center of the target object. 
We also evaluated the performance of a naive model that uses the mean of the LRV values in the training set as the estimate, irrespective
of the image data.

@@

Figure~\ref{fig:case3RFResponse} shows the responses of the first two AMA receptive fields to the training stimuli.
Although the responses vary systematically with target object LRV, there is considerable overlap in the receptive field responses for stimuli having different LRV values.
This overlap is due to the combined effect of the variation in the illumination and background surface spectra.
Recall, however, that the performance shown Figure~\ref{fig:case3Estimates} is based on six rather than two receptive fields;
so that there is likely to be less overlap in the full six-dimensional response space.
That is, inclusion of more receptive fields will in general reduce the ambiguity that is seen when we visualize the responses of just the first two receptive fields.
This effect is shown the right panel of Figure~\ref{fig:RMSEvsRF}, which also illustrates the effect of training set size on estimation performance. 
%% DHB: Check rewrite of last two sentences: OK?
%% VS to add REF to last sentence above.

@@

%% DHB: We need to write something interpretive about the RF structure shown in panel c.

Figure~\ref{fig:case3RFs} shows the first two receptive fields learned for Condition 3.
Similar to Condition 1, the L and M cone excitations receive large weights, while the S cone excitations are largely ignored. 
Additionally, the cone excitations at the target have opposite sign compared to background object locations.
This make sense.
As discussed earlier, luminance is primarily determined by L and M cones, so the S cones should make little
contribution.
Additionally, for this condition, the cone contrast needs to be compared across the target and background regions. 
A center-surround receptive field facilitates such comparison.
However, one should not interpret too much into the features that arise due to the specific geometry of the scene.

%% VS added ->
Our software allows us to compare the effect of background surface reflectances on target object LRV with and without secondary reflections from background objects. 
Our analysis shows that the secondary reflections have minimal effect on LRV estimates.
The estimates without secondary reflections are comparable to that with reflections.
This shows that the primary source of the estimation error is due to the variability in the contrast estimation due to finite number of samples in the background and not due to secondary reflections.

@@

@@

Here we used labeled images rendered from virtual scene descriptions. 
There are many advantages to using rendered images.
One advantage is that they allow us to work with large numbers of labeled images where object reflectance is precisely known at each pixel.
A second advantage is that we can control the variation in distinct scene factors that might affect the difficulty of the estimation problem.
This flexibility allows the study of individual scene factors as well as their combined effects.
Here, we exploited this flexibility to quantify how variation in the shape of the reflectance spectrum of the target object, the spectrum of the illumination, and the reflectance spectrum of the background objects limit LRV estimation.
We believe this basic approach can be extended to include parametric control over the amount of variation of different factors as well as to include variation in additional factors (see below).
%% VS added ->
For example, we could study the case without secondary of the light on the target object from background objects reflections in the scene, and show that such reflections have minimal effect on LRV estimation.
Such analysis are not possible with real images.
%% DHB: Update if we add additional analyses as per some of the comments above.
% % JDB: I don't understand what the the concluding sentence above is saying.

@@

\subsection{Future Directions}
In the work presented here, we studied computational luminance constancy in virtual scenes with naturalistic spectral variation in light sources and in surface reflectance functions. 
% JDB: Vet change to opening sentence w. vijay and david.

It is natural to begin with spectral variation, because this variation is at the heart of what makes luminance constancy a rich computational problem.
In natural scenes, however, there are other sources of variation that add additional richness and make the problem more difficult to solve.
These include variation in non-spectral properties of objects and lighting in the scene, including object texture, material, and shape as well as lighting geometry. 
The methods we developed here may be generalized to study the effects of variation in these factors.
One could incorporate these other sources of variation into the generation of the scenes and again learn estimators from the corresponding labeled images. 
A challenge for this approach will be to thoughtfully control the increase in problem complexity, both to keep compute time feasible and to ensure that it is possible to extract meaningful insight from the results.

>>>>

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
	\centering
	\caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_a.png}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.55 \textwidth}   
        \caption{}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_b.pdf}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{{\bf Color and luminance constancy:} (a)  The light reflected from an object to the eye depends both on the surface reflectance of the object and on the illumination of the object. 
The reflected light also depends on geometric factors, such as the object's shape, pose, and position. 
The human visual system accounts for variations in the reflected light due to object-extrinsic factors and produces a percept of object color that is relatively stable across such variations, an ability called color constancy. 
(b) Images of a sphere under a fixed illuminant.  
Down each column, the sphere's light reflectance value (LRV) varies, but the relative reflectance spectrum (i.e. its spectral shape) is held fixed.
Across each row, the relative reflectance spectrum varies, but the LRV is held fixed.
We cast the problem of computational luminance constancy to be that of estimating the LRV of a target object from the image, across variation in other scene factors. 
Specifically, we study variation in target object relative surface reflectance, variation in illumination spectrum, and variation in the surface reflectance of the other objects in the scene.}
 \end{figure}
%% DHB: Need to decide and then change with search through document: "relative reflectance spectrum" or "shape of reflectance spectrum".
% I prefer relative, because "shape" in a vision paper generally refers to object shape.
% And we found at least one instance where trying to use "shape" led to an awkward construction.

@@

\section*{Methods} \label{Methods}
% Figure 2
\begin{figure}
\centering
\begin{subfigure}[b]{0.22 \textwidth}
        \caption{Library }
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_a.pdf}
        \label{fig:baseSceneLibrary}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \caption{Mill}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_b.pdf}
        \label{fig:baseSceneMill}
    \end{subfigure}    
    ~
    \begin{subfigure}[b]{0.22 \textwidth}
        \caption{Table-Chairs}    
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_c.pdf}
        \label{fig:baseSceneTableChairs}
    \end{subfigure}
    \caption{{\bf Base scenes.} Each subplot shows a rendering of one base scene without additional inserted objects.  The reflectance spectra of the distinct surfaces and the spectral power distributions of the light sources in each scene have been assigned randomly from statistical models of naturally occurring spectra (see text). The images in this figure have been scaled and tone mapped individually for illustrative purposes.}
\label{fig:baseScenes}. \end{figure}
%% DHB: Let's get the scaling and tone mapping used for these under control, so we know what it is.
%% DHB: Maybe also say that the illumination scaling is much smaller than the range used in the study, so that it is possible to see it with common scaling.

\subsection{Overview}
There are four key parts to our methods.  First, we generate a labeled set of training images.  Second, we use a model of the early visual system to compute the responses of the cone photoreceptor mosaic to the labeled images. Third, we apply AMA to the cone responses, as well as to a contrast normalized version of the cone responses, to learn task-optimal receptive fields (RFs). Fourth, we evaluate how well the responses of the RFs may be decoded to achieve luminance constancy.

@@

@@

% Figure 4
\begin{figure}
	\begin{subfigure}{0.18 \textwidth}
    \centering
        \caption{Target position}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_a.pdf}
        \label{fig:targetPositionVariation}
    \end{subfigure}
    ~
	\begin{subfigure}{0.18 \textwidth}
    \centering
        \caption{Target size/Orientation}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_b.pdf}
        \label{fig:targetSizeOrientation}
    \end{subfigure}
~
\centering
	\begin{subfigure}{0.18 \textwidth}
    \centering
        \caption{Target spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_c.pdf}
        \label{fig:targetVariation}
    \end{subfigure}
    ~
    \begin{subfigure}{0.18 \textwidth}
    \centering
        \caption{Illumination spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_d.pdf}
        \label{fig:illuminationVariation}
    \end{subfigure}
          ~  
    \begin{subfigure}{0.18 \textwidth}
    \centering
        \caption{Background spectra}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure4/Figure4_e.pdf}
        \label{fig:backGroundVariation}
    \end{subfigure}

For the primary results, we used the ``{\it Library}'' base scene and a spherical target object.
The library base scene contains 2 area lights. 
We inserted one additional spherical light source into the scene.
The position and size of the inserted object, inserted light source and viewpoint were held fixed across all 
scenes in the database. The geometry is shown in Figure~\ref{fig:3DScene}.
Surface and illuminant spectra were drawn according to statistical models of naturally occurring spectra, as described below.
Multispectral images ($320 \times 240$ pixels) were rendered at 31 evenly-spaced wavelengths between $400$nm and $700$nm.

% Figure 5: Conditions Studied
\begin{figure}
\centering
	\begin{subfigure}{0.33 \textwidth}
		\caption{Condition 1}
%		\includegraphics[width=\textwidth]{../FiguresDraft5/Figure2/Figure2_a.png}
		\includegraphics[width=\textwidth]{../FiguresDraft5/Figure1/Figure1_b.pdf}
 		\label{fig:backgroundVarying}
	\end{subfigure}
	\begin{subfigure}{0.33 \textwidth}
        \caption{Condition 2}	
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure5/Figure5_b.pdf}
        \label{fig:targetIlluminantVarying}
    \end{subfigure}
	\begin{subfigure}{0.33 \textwidth}
	\caption{Condition 3}	
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure5/Figure5_c.pdf}        
        \label{fig:allSpectraVarying}
    \end{subfigure}    
    \caption{{\bf Conditions 1-3:} sRGB renderings of illustrative images for the three types of spectral variations studied in this paper. (a) Condition 1: variable target object relative reflectance spectrum, fixed light source spectra, fixed background object spectra (same as Fig.~\ref{fig:introExampleFigure}). 
(b) Condition 2: variable target object relative reflectance spectrum, variable light source spectra, fixed background object spectra. (c) Condition 3: variable target object relative reflectance spectrum, variable light source spectra, variable background object spectra. The spheres in each row of each panel have the same LRV, while the spheres in each column of each panel have the same relative surface reflectance.  Across panels a-c, spheres in corresponding locations have the same surface reflectance. In all three panels, the overall light source spectra scale factors (see \nameref{Methods}) were drawn from a uniform distribution on the range [0.5, 1.5]. This is smaller than the variation we studied, but allows us to show the variation within the dynamic range available for display in publication. The sRGB values for all three subplots were tone mapped and normalized using a common scale factor prior to gamma correction.} 
\label{fig:studiedCases}
\end{figure}

\subsubsection{Illumination spectra}
To generate illumination spectra, we first developed a statistical model of the Granada daylight measurements (\citeNP{hernandez2001color}; \href{http://colorimaginglab.ugr.es/pages/Data}{http://colorimaginglab.ugr.es/pages/Data}) and then sampled randomly from the model (see appendix).
To match the wavelength spacing we use in rendering, we resampled the wavelength spacing of the Granada spectra to
to 31 evenly-spaced wavelengths between 400 nm and 700 nm.

% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{Granada dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft5/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Statistical model of illumination spectra:} (a) Normalized Granada dataset. Each spectrum is normalized by its mean power. (b) Sample spectra generated using the statistical model derived from normalized Granada dataset. (c) CIE xy chromaticities of the Granda dataset (black) and the samples shown in b (red). (d) sRGB renditions of the samples shown in b.}
\label{fig:illuminant}
\end{figure}

@@

We focussed our analysis on image regions local to the target by cropping the rendered images to $1 \times 1$ degrees of visual angle around the target object ($51 \times 51$ pixels; Figure~\ref{fig:croppedImage}). In foveal primary visual cortex, receptive fields have a spatial extent of approximately 1 degree of visual angle \cite{gattass1981visual, gattass1988visuotopic}. 
The local analysis is thus motivated by the fact that neural receptive fields early in the visual pathways (e.g., retina, primary visual cortex) pool information locally.

@@

\subsubsection*{Learning optimal filters}
AMA is a task-specific Bayesian method for dimensionality reduction.
When provided with a labeled training set, a receptive field response model, a decoder that uses these responses to estimate the stimulus label, and a cost function, AMA returns a set of N linear receptive fields, where N is a parameter of the procedure.
These receptive fields are chosen so that decoding their responses leads to minimum-expected-cost estimation, relative to any other choice of N linear receptive fields.
This is possible because given a training set together with a set of N linear receptive fields and an explicit cost function, it is possible to determine the minimum-expected-cost estimator that maps the receptive field responses to the stimulus labels, on the assumption that all images in the training set are equally likely.
Moreover, the corresponding expected estimation cost may be explicitly computed.
Thus, AMA searches over the space of linear receptive fields to find the set that minimizes the expected estimation cost.
We used the Kullback-Leibler divergence cost function (corresponding to the maximum a posteriori estimator) and mean squared error cost function (corresponding to mean estimator) and assumed that receptive field responses were corrupted by scaled Gaussian noise (i.e. Poisson-like noise with a fano factor of 1.3).
Both, the Kullback-Leibler divergence cost function and mean squared error cost function gave similar results.
The results in this work are for the Kullback-Leibler divergence cost function (see below).
Details of how AMA learns filters and how the filter responses are optimally decoded are provided in previously published work \cite{geisler2009optimal,burge2017accuracy,jaini2017linking}.
For each condition, we learned the filters with a training set consisting of 90\% of the images.

@@

Once the optimal receptive fields have been learned, we need to develop a general decoder that can be applied to arbitrary test stimuli.
A general decoder is necessary because the decoder used to learn the receptive fields requires the response mean and response variance of the filters to every labelled stimulus in the training set \cite{geisler2009optimal,burge2017accuracy}.
To proceed, first we use the AMA optimal receptive fields and the training dataset to find the conditional distributions of the receptive field responses, conditioned on the labelled target object LRVs.
Then, we approximate these with multivariate Gaussian distributions.
To estimate the LRV of the target object in a test image, we use Bayes' rule together with the Gaussian conditional distributions and a uniform prior to obtain the posterior distribution over LRV values given the filter responses.
The optimal estimate is the LRV value that minimizes the expected value of the cost function, where the expectation is taken over the posterior.
Here we again used both Kullback-Leibler divergence cost function and mean square error cost function, so that the optimal estimator is the maximum a posteriori estimate and the mean estimate, respectively.
The performance (relative error) in both cases were comparable, but the results obtained using the mean square error cost function had a bias near the lower and the upper limits of the LRV values used in our simulations.
The bias results because the LRV values are bounded in the range [0.2, 0.6] in estimating the posterior.
For this reason, we present the results for the Kullback-Leibler divergence cost function.

The Gaussian approximations for the posterior were estimated using the same training set that was used to learn the receptive fields.
Estimation performance was evaluated using a test set consisting of the 10\% of images that were not in the training set.
For our primary results, estimation was based on 6 receptive fields.
%% JDB: I thought we were using a squared error cost function. See above?
%% DHB: Wouldn't it make more sense to use MSE and posterior mean for both learning and decoding, since we are reporting relative RMSE as our performance measure?

\subsubsection*{Baseline methods}
To provide baselines for evaluation of the estimation method described above, we used linear regression.
First, we solved for the weights on the average L, M and S cone excitations corresponding to the target object that best predicted the target LRV values.
We took the average cone responses from a $3 \times 3$ region at the center of the target object.
This baseline method disregards information carried by the image regions corresponding to the background objects.
We also performed regression on the contrast normalized version of the cone responses.
This second baseline method indirectly incorporates information carried in the image regions corresponding to the background objects,
through their effect on the contrast normalized responses at the center of the target object. 
We also evaluated the performance of a naive model that uses the mean of the LRV values in the training set as the estimate, irrespective
of the image data.

@@

Figure~\ref{fig:case3RFResponse} shows the responses of the first two AMA receptive fields to the training stimuli.
Although the responses vary systematically with target object LRV, there is considerable overlap in the receptive field responses for stimuli having different LRV values.
This overlap is due to the combined effect of the variation in the illumination and background surface spectra.
Recall, however, that the performance shown Figure~\ref{fig:case3Estimates} is based on six rather than two receptive fields;
so that there is likely to be less overlap in the full six-dimensional response space.
That is, inclusion of more receptive fields will in general reduce the ambiguity that is seen when we visualize the responses of just the first two receptive fields.
This effect is shown the right panel of Figure~\ref{fig:RMSEvsRF}, which also illustrates the effect of training set size on estimation performance. 
%% DHB: Check rewrite of last two sentences: OK?

@@

%% DHB: We need to write something interpretive about the RF structure shown in panel c.
%% VS: added ->
Figure~\ref{fig:case3RFs} shows the first two receptive fields learned for Condition 3.
Similar to Condition 1, the L and M cone excitations receive large weights, while the S cone excitations are largely ignored. 
Additionally, the cone excitations at the target have opposite sign compared to background object locations.
This make sense.
As discussed earlier, luminance is primarily determined by L and M cones, so the S cones should make little
contribution.
Additionally, for this condition, the cone contrast needs to be compared across the target and background regions. 
A center-surround receptive field facilitates such comparison.
However, one should not interpret too much into the features that arise due to the specific geometry of the scene.

@@

%% VS: added ->
Our software allows us to compare the effect of background surface reflectances on target object LRV with and without secondary reflections from background objects. 
Our analysis shows that the secondary reflections have minimal effect on LRV estimates.
The estimates without secondary reflections are comparable to that with reflections.
This shows that the primary source of the estimation error is due to the variability in the contrast estimation due to finite number of samples in the background and not due to secondary reflections.

@@

Here we used labeled images rendered from virtual scene descriptions. 
There are many advantages to using rendered images.
One advantage is that they allow us to work with large numbers of labeled images where object reflectance is precisely known at each pixel.
A second advantage is that we can control the variation in distinct scene factors that might affect the difficulty of the estimation problem.
This flexibility allows the study of individual scene factors as well as their combined effects.
Here, we exploited this flexibility to quantify how variation in the shape of the reflectance spectrum of the target object, the spectrum of the illumination, and the reflectance spectrum of the background objects limit LRV estimation.
We believe this basic approach can be extended to include parametric control over the amount of variation of different factors as well as to include variation in additional factors (see below).
% % JDB: I don't understand what the the concluding sentence above is saying.
%% DHB: Update if we add additional analyses as per some of the comments above.
%% VS added ->
For example, we could study the case without secondary of the light on the target object from background objects reflections in the scene, and show that such reflections have minimal effect on LRV estimation.
Such analysis are not possible with real images.

@@

\subsection{Future Directions}
In the work presented here, we studied computational luminance constancy in virtual scenes with naturalistic spectral variation in light sources and in surface reflectance functions. 
% JDB: Vet change to opening sentence w. vijay and david.
It is natural to begin with spectral variation, because this variation is at the heart of what makes luminance constancy a rich computational problem.
In natural scenes, however, there are other sources of variation that add additional richness and make the problem more difficult to solve.
These include variation in non-spectral properties of objects and lighting in the scene, including object texture, material, and shape as well as lighting geometry. 
The methods we developed here may be generalized to study the effects of variation in these factors.
One could incorporate these other sources of variation into the generation of the scenes and again learn estimators from the corresponding labeled images. 
A challenge for this approach will be to thoughtfully control the increase in problem complexity, both to keep compute time feasible and to ensure that it is possible to extract meaningful insight from the results.
